Stack Safety is a Security Property

### 2021-04-15 tex/main.tex
Reduced section 4, starting on 5

Our stack-safety criteria depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program {\em annotated} with the locations of
instructions representing calls.%, returns, and function entry points\BCP{I
%  still want to talk about whether returns and entry points need to be
%  annotated.  To me, these annotations do not make intuitive sense: it is
%  the caller's responsibility to know whether the address that it is calling
%is really an entry point, and we can trust it to check this.  Similarly, it
%is not the callee's role to say when it is returning; it is the caller's
%prerogative to decide when it has been returned to!  I.e., from the caller's
%point of view, it has been returned to when control resumes at the
%instruction following the call with the SP reset to where it was at the
%point of the call. (This is the point where it wants to be able to access
%its own local state again and expected to find it unmolested.)}.
Although these
annotations would typically come from a compiler together with the machine code,
we do not assume anything about their provenance.

@@

\BCP{I believe we could delete this section---and never mention WBCF at all
  (except to say that we leave it out, compared to STkTokens)---if we
  eliminated return (and entry point) annotations.}

Since we rely heavily on code annotations, we must ask: what if the annotations
are wrong? First, suppose a call or return occurs that is not annotated.
In this case, the answer is simple: as far as the properties are concerned, there has
been no call or return, only an instruction sequence resembling such an operation, and
therefore no associated change in permissions. This is particularly interesting in the
case of a false return, where control may appear to have returned to the caller, but
the enforcement mechanism must still protect the caller's memory.

Alternatively, an annotated instruction might not really be a call or return.
In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
and returns are implemented by fixed sequences of instructions which the tag policy
will force to be executed from start to finish. The annotated call or return instruction should always
be the final instruction of the appropriate sequence. What if an annotated instruction exists outside
such a sequence? We will statically rule out some scenarios with a well-formedness criterion
on initial states: annotated instructions will initially be tagged apropriately, as the
final instruction in the call or return sequence, and all blessed\bcp{undefined} sequences will end with
annotated instructions. This is all we need for the policy to
protect itself: it can then failstop if it would execute the blessed sequence out of order.
So, even in our strong attacker model, we need very little in the way of restrictions on
the code when considering micropolicy-based enforcement.  \BCP{Is this
  discussion still needed?  I had hoped that switching to our ``new''
  definition on subtraces would eliminate such complexities!}

\BCP{Not sure what this is about.}In general, regardless of enforcement mechanism, we take this approach:
a given enforcement mechanism may place a static well-formedness criterion
on the code it applies to. Weaker enforcement mechanisms might also need dynamic
reasonableness conditions to prevent the defender from sabotaging enforcement, as
discussed in section \ref{sec:cheri}.  \BCP{Weak.}

@@

This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
We believe that it is flexible enough to capture code-altering approaches, such as code
rewriting or capability-based techniques, though perhaps with less elegance.

@@

%The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
%at the time of the call. The stack can be used to pass arguments or return results
%in the usual way, by placing them at the top of the caller's frame. This will be modeled
%as a form of sharing in section \sna{cite the sharing section}.
%Callers and callees can also communicate via registers if they wish; indeed, our
%definitions say nothing at all about registers (except $\PCname$ and $\rsp$).

%\iftrue\bcp{This implicitly amounts to choosing a caller-saves register
%  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi

>>>>

Our stack-safety criteria depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program {\em annotated} with the locations of
instructions representing calls, and we determine when a return has occurred by tracking
control flow and observing when we reach a {\em return target}. For a given state annotated
as a call, in a standard calling convention, another state fulfills its return target if
it has the same stack pointer and the program counter is at the instruction following
the call, but we abstract over different conventions.
%, returns, and function entry points\BCP{I
%  still want to talk about whether returns and entry points need to be
%  annotated.  To me, these annotations do not make intuitive sense: it is
%  the caller's responsibility to know whether the address that it is calling
%is really an entry point, and we can trust it to check this.  Similarly, it
%is not the callee's role to say when it is returning; it is the caller's
%prerogative to decide when it has been returned to!  I.e., from the caller's
%point of view, it has been returned to when control resumes at the
%instruction following the call with the SP reset to where it was at the
%point of the call. (This is the point where it wants to be able to access
%its own local state again and expected to find it unmolested.)}.
Although call annotations would typically come from a compiler together with
the machine code, we do not assume anything about their provenance.

@@

\BCP{I believe we could delete this section---and never mention WBCF at all
  (except to say that we leave it out, compared to STkTokens)---if we
  eliminated return (and entry point) annotations.}
\sna{I'm not sure we can escape questions about annotations, even if we have
  just calls, but I tried to abbreviate it.}

Since we rely heavily on call annotations, we must ask: what if the annotations
are wrong? First, suppose a call occurs that is not annotated.
In this case, the answer is simple: as far as the properties are concerned, there has
been no call, only an instruction sequence resembling one, and
therefore no associated change in permissions. Then the caller is guaranteed no
protections.

Alternatively, an annotated instruction might not obey the calling convention.
In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
and returns are implemented by fixed sequences of instructions which the tag policy
will force to be executed from start to finish, termed {\em blessed sequences}.
When applying our properties to a micro-policy setting, we will statically rule out some
scenarios with a well-formedness criterion on initial states: all annotated instructions will
be situated at the appropriate instruction of a blessed sequence, tagged according to the policy.
This is all we need for the policy to protect itself: it can then failstop if it would
execute the blessed sequence out of order.

@@

%This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
%We believe that it is flexible enough to capture code-altering approaches, such as code
%rewriting or capability-based techniques, though perhaps with less elegance.

@@

Unless otherwise noted, colored boxes indicate an assignment of the address to a
{\em domain} -- a label reflecting a particular set of access rights. Domains
are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
(outside the stack entirely.)
